---
title: "Class Reference"
author: "Student name"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

## Installed Packages

>tidyverse

## Jan. 27

Set working directory = setwd("#").
Code block = ``` and then {r} and then ```.
Load a library = library(#).

```{r}
setwd("~/GitHub/data_journalism_2022_spring")
install.packages("tidyverse")
library(tidyverse)
```

## Feb. 1

arrange in ascending order = arrange().
arrange in descending order = arrange(desc()).
setting a variable = var <- code #preferable
setting a variable = code -> var 
  ex:
  
```{r}
  var <- ppp_maryland_loans %>%
  summarise(
    count_loans = n(),
    total_loans_amount = sum(amount)
  ) %>%
  arrange(total_loans_amount)
  
```

add/replace/change a column = mutate(column_name = value/object).
value = number.
value = 'word'.
make all values in a column uppercase = str_to_upper(column_name).
show me specific columns A, B, and C = select(column_A, column_B, column_C).

## Feb. 3

Finding all entries of a certain summarized column = view().
  ex:

```{r}
maryland_ppp %>%
  group_by(project_county_name) %>%
  summarize(
    counties = n()
  )
view(counties)

```

Changing a data type to <date> = ymd_hms() OR ymd().
  year month date = yms.
  hour minute second = hms.
  If there's no time, it's just yms().
  ex:
  
```{r}
lubridate_maryland_ppp <- maryland_ppp_sample %>% mutate(
  CleanDate = ymd_hms(date_approved)
)
glimpse(lubridate_maryland_ppp)

```

