---
title: "Class Reference"
author: "Student name"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

## Installed Packages

>tidyverse

## Jan. 27

Set working directory = setwd("#").
Code block = ``` and then {r} and then ```.
Load a library = library(#).

```{r}
setwd("~/Github/data_journalism_2022_spring")
install.packages("tidyverse")
library(tidyverse)
```

## Feb. 1

arrange in ascending order = arrange().
arrange in descending order = arrange(desc()).
setting a variable = var <- code
  ex:
  
```{r}
  var <- ppp_maryland_loans %>%
  summarise(
    count_loans = n(),
    total_loans_amount = sum(amount)
  ) %>%
  arrange(total_loans_amount)
  
```

