---
title: "lab_08"
author: "derek willis"
date: "8/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Census API key

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and tidycensus.

**Task** load these two packages

```{r}
# Turn off scientific notation
options(scipen=999)
library(tidyverse)
library(tidycensus)
```

## Setup Census API

You'll need your Census API Key:

**Task** set your API Key

```{r echo=FALSE}
census_api_key("102c0c72d8f8eb880608f9dacb53fdf9100f8ba0")
```

## Load and view ACS variables

You'll also need to load the list of American Community Survey variables from the 2019 5-year estimates so you can identify the codes you need:

**Task** load the variables from the 2019 5-year ACS and save them to a dataframe

```{r}
# Write code here to load the ACS variables
acs5 <- load_variables(2019, "acs5", cache = TRUE)
```

## Load PPP data

You'll also need to load an RDS file of PPP loan applications that were removed from the original data by lenders in July.

**Task** read the RDS file and save it to a dataframe

```{r}
#Write code here to read the RDS file
ppp_applications <- read_rds("data/lab_08.rds")
```

## Answer questions

**Q1** What is the Maryland jurisdiction with the lowest median age, according to the ACS 2019 5-year estimates? You will need to find the variable for median age first in your list of ACS variables.
**A1** The lowest median age is 35.4 years, in Baltimore City.

```{r}
get_acs(
  geography = "county",
  variables = c(medage = "B01002_001"),
  state = "MD", 
  year = 2019
) %>% 

arrange(estimate)

```

**Q2** Which Maryland counties have a median age of at least 50? Write code to isolate just those jurisdictions.
**A2** Worcester and Talbot counties.

```{r}
get_acs(
  geography = "county",
  variables = c(medage = "B06002_001"),
  state = "MD", 
  year = 2019
) %>% 
  
filter(estimate > 50) %>% 

arrange(estimate)

```

**Q3** Which Maryland counties have the highest percentage of 5-17 year-olds that speak Spanish at home? You'll need one dataframe with county-level data on the number of 5-17 year-olds that speak Spanish at home and a second dataframe with the total population of 5-17 year-olds by county. For each dataframe, rename the `estimate` column to a meaningful name you've given the ACS variable, remove the `variable` columns in each using select and then join them based on GEOID and NAME. Then you can add a column to calculate the percentage of the total 5-17 population that speaks Spanish at home.
**A3** Prince George's County has by far the highest percentage of Spanish-speaking 5- to 17-year-olds.

```{r}
spanish_speaking <- get_acs(
  geography = "county",
  state = "MD",
  year = 2019,
  variables = c(kids_spanish = "B16003_004")
) %>% 
  select(-variable)

kids_total <- get_acs(
  geography = "county",
  state = "MD",
  year = 2019,
  variables = c(kids_total = "B16007_002")
) %>% 
  select(-variable)


kids_total %>% 
  inner_join(spanish_speaking, by=c("GEOID", "NAME")) %>% 
  rename(total=estimate.x, spanish=estimate.y, moe_total=moe.x, moe_spanish=moe.y) %>% 
  mutate(percent = spanish/total*100) %>% 
  arrange(desc(percent))

```

**Q4** Given the counties with the largest percentage of 5-17 year-olds that speak Spanish at home, what stands out to you about the data? How would you describe the top 5 counties? How would you describe Garrett County's numbers?
**A4** Prince George's County has by far more Spanish-speaking children than the rest of the top 5 counties. Garrett County has no Spanish-speaking children, but is listed as having only about 4000 children. The margin of error for that county is 22, so it very well could have Spanish-speaking children.

**Q5** Which 5-digit zip code had the most removed PPP applications, and what is the percentage of Black population in that zip code according to the ACS? To answer this question, you will need to create separate dataframes with ACS data on total population by zip code and Black population by zip code. In each dataframe, rename the `estimate` column and remove the `variable` column. You'll need to combine the two dataframes in order to calculate the percentage, and then you can combine it with the PPP data. Remember: the PPP data may not have only 5-digit zip codes!
**A5** The zip code with the most removed applications  is 60620. 95.7% of people are Black in this zip code.

```{r}
ppp_applications <- ppp_applications %>%
  mutate(zip5 = str_sub(zip, 1, 5))

total_pop <- get_acs(
  geography = "zcta",
  year = 2019,
  variables = c(pop_total = "B01001_001")
)

black_pop <- get_acs(
  geography = "zcta",
  year = 2019,
  variables = c(black_pop = "B02001_003")
)

total_pop <- total_pop %>% 
  inner_join(black_pop, by=c("GEOID")) %>% 
  rename(total=estimate.x, black=estimate.y, moe_total=moe.x, moe_black=moe.y, zip5=GEOID) %>% 
  mutate(percent = black/total*100) %>% 
  select(-variable.x, -variable.y, -NAME.x, -NAME.y)

total_pop_ppp <- total_pop %>% 
  left_join(ppp_applications, by=c("zip5"))

total_pop_ppp %>% 
  group_by(zip5, percent, city) %>% 
  summarize(
    count = n()
  ) %>% 

arrange(desc(count))


```

**Q6** How would you describe to an editor the top 10 zip codes with the most removed applications in terms of Black population? Where are they located, and what do they have in common? Are there any zip codes that seem different, and if so why?
**A6** The top 10 zip codes are almost all in Chicago and almost all in Black majority zip codes. There are two outliers: A zip code in Atlanta, although it is also majority Black, and a zip code in Chicago that is not majority Black. The differences might be because the non-majority black zip code in Chicago is close to the other zip codes and therefore has a similar business and labor experience.

**Q7** What other columns from the PPP data could be useful in reporting on what happened to these removed loans?
**A7** The lender and the amount.
