---
title: "lab_01"
author: "sean mussenden"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the slice of PPP loans for West Virginia.

**Task** Run the codeblock below to load the data.

```{r}
wv_ppp_loans <- read_rds("lab_01.rds")
```

## Examine Data

**Task** Run the codeblock below to use glimpse to get a sense of the column names, types and sample data.
**Q1** How many rows and columns does this dataframe have?
**A1** [33,336 rows, 58 columns]

```{r}
glimpse(wv_ppp_loans)
```

## Answer questions

Use the pre-lab and your reference notebook as examples.

**Q2** Which West Virginia county got the most loans? How many?
**A2** [Kanawha County has 4273 loans.]

```{r}
# Write the code to produce your answer here
wv_ppp_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n()
  ) %>%
  arrange(desc(count_loans))

```
**Q3** Which West Virginia county got the fewest loans? How many?
**A3** [The West Virginia county with the fewest loans, Wirt County, has 36 loans. The county with the fewest loans overall is Palm Beach County, with 2 loans.]

```{r}
# Write the code to produce your answer here
wv_ppp_loans %>%
  group_by(project_county_name, project_state) %>%
  summarise(
    count_loans = n()
  ) %>%
  arrange(count_loans)

```
**Q4** What's your best educated guess to explain why the county identified in question 2 had a lot of loans and why the county identified in question 3 had only a few loans? Hint: you won't find the answer in the data, you'll need to do some quick web research. You do not need to write any code for this question, so no codeblock is provided.  
**A4** [Kanawha is the most populous county in West Virginia, whereas Palm Beach County is not a West Virginia county -- it's in Florida.]


**Q5** What West Virginia business got the state's largest PPP loan? How much was it for? What does the business do? How many jobs did the loan retain, according to the data?
**A5** [Greenbriar Hotel Corporation got the state's largest PPP loan. It was for $8,912,202. The business is a hotel corporation. According to the data, the loan retained 500 jobs.]

```{r}
# Write the code to produce your answer here
wv_ppp_loans %>%
  group_by(name, amount, jobs_retained) %>%
  summarise(
    total_loan_amount = sum(amount)
  ) %>%
  arrange(desc(total_loan_amount))

```

**Q6** What West Virginia politician (with his family) owns the company identified in question 5? What's your source for that information? Hint: this will require web research, you won't be able to answer this purely from the data set.  You do not need to write any code for this question, so no codeblock is provided.  
**A6** [Jim Justice and his family own the company. My source is the Office of the Governor of West Virginia: https://governor.wv.gov/about/Pages/default.aspx]

**Q7**. Produce a summary table that has the following information for West Virginia as a whole (the whole state, not by county):
* count of loans
* total amount of all loans
* the average loan amount
* the median loan amount
* the smallest loan amount
* the largest loan amount

Store it as a variable object called `wv_summary` so that you can examine it with the data viewer. What newsworthy nuggets jump out at you that you could imagine dropping into a news story? Write a few summary sentences with overall findings.

**A7** [write your answer here]
```{r}
# Write the code to produce your answer here
wv_ppp_loans %>%
  summarise (
      count_loans = n(),
      total_loan_amount = sum(amount),
      average_loan_amount = mean(amount),
      median_loan_amount = median(amount),
      min_loan_amount = min(amount),
      max_loan_amount = max(amount)
    )

# I couldn't figure out this question.
  
```